Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-06-19 articles/formalism/sections/kernel_paintbrush.tex
Editing text and commenting through generalized E/B

\section{Understanding polarization signatures of magnetized filaments}
% 
\begin{figure}[t]
\includegraphics[width=0.5\columnwidth]{line.pdf}
\includegraphics[width=0.5\columnwidth]{spiral.pdf}
\includegraphics[width=1.0\columnwidth]{colorbar.pdf}
\caption{ The polarization signals of toy filament structures. In a filament organized perfectly along a magnetic field line, the polarization will be perpendicular to the filament direction.  The $E/B$ modes of filaments are in some ways easier to think about than the Stokes parameters. Left panels: in a straight filament, the E-mode is positive along the filament and at the ends, but negative along the sides.  B-modes are only non-zero at the ends.  Right panels: in a curved filament, the E-mode is again positive along the filament.  Outside the filament, the $E$-mode is more negative on the interior of the curve than the exterior.  The $B$-modes are again non-zero only at the ends, and are akin to the straight filament case. In all images, the longitude angle increases to the left (East in sky convention).  All plot are on a common, arbitrary color scale.}
\label{fig:polfilaments}
\end{figure}
%

The real space kernels give us a better intuitive understanding of the $E/B$ modes associated with physical objects.  For example, a simple model for a magnetize filament has the magnetic field threaded along a linear gas overdensity.  Precession of the dust grains around the magnetic field leads to a net polarization perpendicular to the magnetic field (and perpendicular to the filament overall).  For a filament aligned North--South, the polarization will be horizontal or $Q<0$, $U=0$ (left pane of \fig{fig:polfilaments}).  The Green's function kernels for horizontal polarization are rotated by 90 degrees relative to the components of $\cal{M}_G$ in \fig{fig:vis_kernel}.

@@

The non-zero $B$ result is somewhat surprising given that the polarization pattern is symmetric to both horizontal and vertical reflections through the filament center.  However, unlike a circular ring, this filament is not a configuration with a definite parity.  Since the scalar description of polarization is coordinate independent, the $E/B$ patterns do not depend on the orientation of the filament.  A filament inclined at $45^\circ$ will have a similar $E/B$ pattern, but different reflection symmetries.

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern. Here it is important to note that this cancellation will not happen for an ellipse or any loop of non-contant radius of curvature.  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside. While the $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations, note that it is non-vanishing in the intermediate regions due to the varying radius of curvature of the spiral \footnote{ \revisit{Why do we use this specific spiral - Its easy to compute the tangent to the spiral at any point.}}. \comment{If this is correct, then one expects to see a higher B-mode signal in intermediate regions of the spiral, as compared to the B-mode signal in the intermediate regions of the linear filament. This  then should be more visible in a logarithmically scaled version of Fig. 8.}

>>>>

\section{Understanding polarization signatures of magnetized filaments}
% 
\begin{figure}[t]
\includegraphics[width=0.5\columnwidth]{line.pdf}
\includegraphics[width=0.5\columnwidth]{spiral.pdf}
\includegraphics[width=1.0\columnwidth]{colorbar.pdf}
\caption{ The polarization signals of toy filament structures. In a filament organized perfectly along a magnetic field line, the polarization will be perpendicular to the filament direction.  The $E/B$ modes of filaments are in some ways easier to think about than the Stokes parameters. Left panels: in a straight filament, the E-mode is positive along the filament and at the ends, but negative along the sides.  B-modes are only non-zero at the ends.  Right panels: in a curved filament, the E-mode is again positive along the filament.  Outside the filament, the $E$-mode is more negative on the interior of the curve than the exterior.  The $B$-modes are again non-zero only at the ends, and are akin to the straight filament case. In all images, the longitude angle increases to the left (which is East in sky convention).  All plot are on a common, arbitrary color scale.}
\label{fig:polfilaments}
\end{figure}
%

The real space kernels give us a better intuitive understanding of the $E/B$ modes associated with physical objects.  For example, a simple model for a magnetized filament has the magnetic field threaded along a linear gas overdensity.  Precession of the dust grains around the magnetic field leads to a net polarization perpendicular to the magnetic field (and perpendicular to the filament overall).  For a filament aligned North--South, the polarization will be horizontal or $Q<0$, $U=0$ (left pane of \fig{fig:polfilaments}).  The Green's function kernels for horizontal polarization are rotated by 90 degrees relative to the components of $\cal{M}_G$ in \fig{fig:vis_kernel}.

@@

The non-zero $B$ result is somewhat surprising given that the polarization pattern is symmetric to both horizontal and vertical reflections through the filament center.  However, unlike a circular ring, this filament is not a configuration with a definite parity.  Since the scalar description of polarization is coordinate independent, the $E/B$ patterns do not depend on the orientation of the filament.  A filament inclined at $45^\circ$ will have a similarly inclined $E/B$ pattern, but different reflection symmetries.

The intuition from the real-space kernels holds also when we distort the shape of the filament.  If the filament were bent around into a circle, the positive and negative parts of the $B$ pattern will cancel, and we are left with a hoop of pure $E$ pattern. Here it is important to note that this cancellation will not happen for an ellipse or any loop of non-contant radius of curvature.  The same general description holds for a spiral-shaped filament, which can can be viewed as distortion of the straight filament.  The filament is highlighted by positive $E>0$.  The $E$-pattern is more negative on the interior of a curve than on the exterior, and the concentric rings of filamentary structure make an increasingly negative $E$ value inside. While the $B$-pattern is again concentrated at the ends of the filament in an oriented pair of positive/negative fluctuations, note that it is non-vanishing in the intermediate regions due to the varying radius of curvature of the spiral.
%\footnote{ \revisit{Why do we use this specific spiral - Its easy to compute the tangent to the spiral at any point.}}.

\comment{If this is correct, then one expects to see a higher B-mode signal in intermediate regions of the spiral, as compared to the B-mode signal in the intermediate regions of the linear filament. This  then should be more visible in a logarithmically scaled version of Fig. 8.}

### 2018-06-19 articles/formalism/sections/real_space_formalism.tex
Editing text and commenting through generalized E/B

\subsection{Matrix notation} \label{sec:mat_pol_intro}
Our derivation of real space operators is more transparent in a matrix-vector notation.\footnote{While we work with the matrix and vector sizes given in terms of some pixelization parameter $\rm N_{\rm pix}$, all the relations are equally valid in the continuum limit attained by allowing $\rm N_{\rm pix}\rightarrow \infty$}
We introduce a matrix that encodes spin spherical harmonic basis vectors:
%
\beq
{}_{|s|}\mathcal{Y}= \bmat _{+s}Y & 0 \\ 0 & _{-s}Y \emat _{2 \rm N_{\rm pix} \times 2 \rm N_{\rm alms}} ;\qquad  {}_{|s|}\mathcal{Y}^{\ddagger}= \Delta \Omega \bmat _{+s}Y^{\dagger} & 0 \\ 0 & _{-s}Y^{\dagger} \emat _{2 \rm N_{\rm alms} \times 2 \rm N_{\rm pix}} \,,
\eeq
%
where $s$ denotes the spin of the basis functions and our \revisit{definition of ${}_{|s|}\mathcal{Y}^{\ddagger}$ differs from the conventional conjugate transpose operation by the factor $\Delta \Omega$.  We introduce this to ensure the ortho-\textit{normality} of these operations, so that ${}_{|s|}\mathcal{Y}^{\ddagger} {}_{|s|}\mathcal{Y} = I_{2 \rm N_{\rm alms} \times2 \rm N_{\rm alms}}$ on the discretized sphere, and maintain the standard definition of spherical harmonics.} \comment{[Orthonormal only in infimetesimal pixel limit?, I think it doesnt have to be infitesimal but sufficiently small.]}

@@

where all the symbols have their usual meaning and the explicit multipole dependence of the real space operators has been suppressed for brevity. \revisit{Note that when the operators are expressed in terms of the Euler angles $(\alpha_{qe},\beta_{qe},\gamma_{qe})$ they can be interpreted as the Greens functions and  we denote them by $\mathcal{I}_G=\Delta \Omega \mathcal{I}^*$ and $\mathcal{D}_G=\Delta \Omega \mathcal{D}$. When expressed as function of Euler angles $(\alpha_{eq},\beta_{eq},\gamma_{eq})$ corresponding to the inverse rotations they can be interpreted as some convolving beam and we denote them by $\mathcal{I}_B=\Delta \Omega \mathcal{I}$ and $\mathcal{D}_B=\Delta \Omega \mathcal{D}$. [K: question about $\Delta \Omega$ here.]} Note that unlike in the case of the operators $\mm_G$ and $\mm_B$ which have different shapes owing to their dependence on Euler angles $\alpha$ and $\gamma$ respectively, the operators $D_G$ and $D_B$ are identical since $(\alpha_{qe}-\gamma_{qe}) = (\alpha_{eq}-\gamma_{eq})$, while $\mathcal{I}_{G}$ and $\mathcal{I}_B$ are related by conjugation since  $(\alpha_{qe}+\gamma_{qe}) = -(\alpha_{eq}+\gamma_{eq})$.

>>>>

\subsection{Matrix notation} \label{sec:mat_pol_intro}
Our derivation of real space operators is more transparent in a matrix-vector notation.\footnote{While we work with the matrix and vector sizes given in terms of some pixelization parameter $\rm N_{\rm pix}$, all the relations are equally valid in the continuum limit attained by allowing $\rm N_{\rm pix}\rightarrow \infty$}
We introduce a matrix that encodes spin spherical harmonic basis vectors:
%
\beq
{}_{|s|}\mathcal{Y}= \bmat _{+s}Y & 0 \\ 0 & _{-s}Y \emat _{2 \rm N_{\rm pix} \times 2 \rm N_{\rm alms}} ;\qquad  {}_{|s|}\mathcal{Y}^{\ddagger}= \Delta \Omega \bmat _{+s}Y^{\dagger} & 0 \\ 0 & _{-s}Y^{\dagger} \emat _{2 \rm N_{\rm alms} \times 2 \rm N_{\rm pix}} \,,
\eeq
%
where $s$ denotes the spin of the basis functions and our definition of ${}_{|s|}\mathcal{Y}^{\ddagger}$ differs from the conventional conjugate transpose operation by the factor $\Delta \Omega$.  We introduce this to ensure the orthonormality of these operations on the discretized sphere when the pixel size is sufficiently small, ${}_{|s|}\mathcal{Y}^{\ddagger} {}_{|s|}\mathcal{Y} = I_{2 \rm N_{\rm alms} \times2 \rm N_{\rm alms}}$, and also maintain the standard definition of spherical harmonics.

@@

where all the symbols have their usual meaning and the explicit multipole dependence of the real space operators has been suppressed for brevity. \revisit{Note that when the operators are expressed in terms of the Euler angles $(\alpha_{qe},\beta_{qe},\gamma_{qe})$ they can be interpreted as the Greens functions and  we denote them by $\mathcal{I}_G=\Delta \Omega \mathcal{I}^*$ and $\mathcal{D}_G=\Delta \Omega \mathcal{D}$. When expressed as function of Euler angles $(\alpha_{eq},\beta_{eq},\gamma_{eq})$ corresponding to the inverse rotations they can be interpreted as some convolving beam and we denote them by $\mathcal{I}_B=\Delta \Omega \mathcal{I}$ and $\mathcal{D}_B=\Delta \Omega \mathcal{D}$.} \comment{[K: question about $\Delta \Omega$ and sums over pixels here and in the rest of the kernels.]} Note that unlike in the case of the operators $\mm_G$ and $\mm_B$ which have different shapes owing to their dependence on Euler angles $\alpha$ and $\gamma$ respectively, the operators $D_G$ and $D_B$ are identical since $(\alpha_{qe}-\gamma_{qe}) = (\alpha_{eq}-\gamma_{eq})$, while $\mathcal{I}_{G}$ and $\mathcal{I}_B$ are related by conjugation since  $(\alpha_{qe}+\gamma_{qe}) = -(\alpha_{eq}+\gamma_{eq})$.
